---
title: "Project Stuff"
author: "Jon Matkin"
date: '2023-02-28'
output: html_document
---

```{r}
lg_data <- read.csv('2020_JP_TS_Level4.csv')
head(lg_data)
```
```{r}
LG_mar_apr_data <- read.csv("JP_mar_apr_data.csv")
```

```{r}
precip_data <- read.csv("nasa_precip_data_mar_apr.csv")
head(precip_data)
```
```{r}
nw_precip_data <- read.csv("North West Precip Data.csv")
ne_precip_data <- read.csv("North East Precip Data.csv")
sw_precip_data <- read.csv("South West Precip data.csv")
se_precip_data <- read.csv("South East Precip Data.csv")
```

```{r}
head(nw_precip_data)
head(ne_precip_data)
head(sw_precip_data)
head(se_precip_data)
```

```{r}
hist(nw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf, main="Northwest Region Distribution", xlab="Precipitation (kg m-2)")
hist(ne_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf, main="Northeast Region Distribution", xlab="Precipitation (kg m-2)")
hist(sw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf, main="Southwest Region Distribution", xlab="Precipitation (kg m-2)")
hist(se_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf, main="Southeast Region Distribution", xlab="Precipitation (kg m-2)")
```
Unsurprisingly the data looks very similar for each quadrant

```{r}
sw_precip_scaled <- scale(sw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf)
sw_precip_scaled <- rnorm(sw_precip_scaled)
se_precip_scaled <- scale(se_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf)
se_precip_scaled <- rnorm(se_precip_scaled)
nw_precip_scaled <- scale(nw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf)
nw_precip_scaled <- rnorm(nw_precip_scaled)
ne_precip_scaled <- scale(ne_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf)
ne_precip_scaled <- rnorm(ne_precip_scaled)

scaled_precip  <- data.frame(sw = sw_precip_scaled,
                             se = se_precip_scaled,
                             nw = nw_precip_scaled,
                             ne = ne_precip_scaled)

sw_se <- t.test(scaled_precip$sw - scaled_precip$se, mu=0)
sw_nw <- t.test(scaled_precip$sw - scaled_precip$nw, mu=0)
sw_ne <- t.test(scaled_precip$sw - scaled_precip$ne, mu=0)
se_nw <- t.test(scaled_precip$se - scaled_precip$nw, mu=0)
se_ne <- t.test(scaled_precip$se - scaled_precip$ne, mu=0)
nw_ne <- t.test(scaled_precip$nw - scaled_precip$ne, mu=0)

sw_se
sw_nw
sw_ne
se_nw
se_ne
nw_ne

```


I think it would be better to split the data into each site and do analysis from there
Then we may be able to draw conclusions separately by site. Kind of ties into how I 
want to look at location

```{r}
unique(LG_mar_apr_data$Site_Name)
```
Set up the dataframes based on location

```{r}
library(dplyr)
lg_east <- LG_mar_apr_data %>% filter(Site_Name == "TS_East")
lg_english <- LG_mar_apr_data %>% filter(Site_Name == "TS_English")
lg_finkle <- LG_mar_apr_data %>% filter(Site_Name == "TS_Finkle")
lg_hague <- LG_mar_apr_data %>% filter(Site_Name == "TS_Hague")
lg_indian <- LG_mar_apr_data %>% filter(Site_Name == "TS_Indian")
lg_northwestbay <- LG_mar_apr_data %>% filter(Site_Name == "TS_NorthwestBay")
lg_outlet <- LG_mar_apr_data %>% filter(Site_Name == "TS_Outlet")
lg_polehill <- LG_mar_apr_data %>% filter(Site_Name == "TS_PoleHill")
lg_shlevingrock <- LG_mar_apr_data %>% filter(Site_Name == "TS_ShelvingRock")
lg_sucker <- LG_mar_apr_data %>% filter(Site_Name == "TS_Sucker")
lg_sunset <- LG_mar_apr_data %>% filter(Site_Name == "TS_Sunset")
lg_west <- LG_mar_apr_data %>% filter(Site_Name == "TS_West")
# Need sunset and west
```

Given the scope for this project and prior examination of sites, we will start by using only 3:
East
Outlet
West

Here is some exploratory analysis on those sensors
```{r}
head(lg_east)
#str(lg_east)
dim(lg_east)
```
```{r}
# NA analysis
# Want to drop cols with all values of NA

#summary(lg_east)
cols_drop_east <- c("Turbidity_Water_EXO","Stage_Water_WaterLog","Temperature_Water_WaterLog","Chla_Water_EXO","Flow_Water_SL","Level_Water_SL","PC_Water_EXO","Temperature_Water_SL","Conductivity.46cm_Soil_HydraProbe","Moisture.46cm_Soil_HydraProbe","Temperature.46cm_Soil_HydraProbe","Flow_Water_IQ","Stage_Water_IQ","Temperature_Water_IQ","Time","X","X.1")

east <- lg_east[,!(names(lg_east) %in% cols_drop_east)]
head(east)
```
Now need to do the same thing for outlet and west

```{r}
#summary(lg_outlet)

drop_cols_outlet <- c("Conductivity_Soil_HydraProbe","Moisture_Soil_HydraProbe","Stage_Water_CS451","Temperature_Soil_HydraProbe","Temperature_Water_CS451","Stage_Water_WaterLog","Temperature_Water_WaterLog","Conductivity.46cm_Soil_HydraProbe","Moisture.46cm_Soil_HydraProbe","Temperature.46cm_Soil_HydraProbe","Flow_Water_IQ","Stage_Water_IQ","Temperature_Water_IQ", "X","X.1","Time")

outlet <- lg_outlet[,!(names(lg_outlet) %in% drop_cols_outlet)]
head(outlet)
```

```{r}
#summary(lg_west)

drop_cols_west <- c("Stage_Water_CS451","Temperature_Water_CS451","Chla_Water_EXO","Flow_Water_SL","Level_Water_SL","PC_Water_EXO","Temperature_Water_SL","Conductivity.46cm_Soil_HydraProbe","Moisture.46cm_Soil_HydraProbe","Temperature.46cm_Soil_HydraProbe","Time","X","X.1")

west <- lg_west[,!(names(lg_west) %in% drop_cols_west)]
head(west)
```

Some exploratory analysis of the variables we are interested in
```{r}
boxplot(west$Chloride_Water_EXO, main="Boxplot of Chlorides for West")
boxplot(west$pH_Water_EXO, main="Boxplot of pH for West")
boxplot(west$Temperature_Water_EXO, main="Boxplot of Water Temp for West")
```
```{r}
boxplot(east$Chloride_Water_EXO, main = "Boxplot of chlorides for East")
boxplot(east$pH_Water_EXO, main="Boxplot of pH for East")
boxplot(east$Temperature_Water_EXO, main="Boxplot of Water Temp for East")
```
```{r}
boxplot(outlet$Chla_Water_EXO,main="Boxplot of Chlorides for Outlet")
boxplot(outlet$pH_Water_EXO, main="Boxplot of pH for Outlet")
boxplot(outlet$Temperature_Water_EXO, main="Boxplot of Water Temp for Outlet")
```

Now we can look into the nasa precipitation data

```{r}
tail(precip_data)
```

Have a small problem where the datasets are not exactly the same size for some reason.
West and outlet are only missing one row, but west is missing 20 rows
  Fixed this issue in excel by adding some NA rows
  
  
Need to add the precipitation values from the four different areas in the bounding box
Need to do this for all three sensors
```{r}
east$sw_precip <- sw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
east$se_precip <- se_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
east$nw_precip <- nw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
east$ne_precip <- ne_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
```

```{r}
outlet$sw_precip <- sw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
outlet$se_precip <- se_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
outlet$nw_precip <- nw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
outlet$ne_precip <- ne_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
```

```{r}
west$sw_precip <- sw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
west$se_precip <- se_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
west$nw_precip <- nw_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
west$ne_precip <- ne_precip_data$mean_NLDAS_FORA0125_H_2_0_Rainf
```


Going to export the east, outlet, west to csv so I can use them in python
```{r}
write.csv(east, file="C:\\Users\\matkij\\Documents\\ITWS 6600 Data Analytics\\east.csv")
write.csv(outlet,file= "C:\\Users\\matkij\\Documents\\ITWS 6600 Data Analytics\\outlet.csv")
write.csv(west, file="C:\\Users\\matkij\\Documents\\ITWS 6600 Data Analytics\\west.csv")
```

Now we can start to construct regression models to predict lake health based on 
  rain in the different quadrants
  
```{r}
# Linear regression for chlorides in east
east_chlr_lm <- lm(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=east)
summary(east_chlr_lm)
```
```{r}
# Linear regression for water pH in east
east_ph_lm <- lm(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=east)
summary(east_ph_lm)
```
```{r}
# Linear regression for water temperature in east
east_temp_lm <- lm(Temperature_Water_EXO~ sw_precip + se_precip + nw_precip + ne_precip,data=east)
summary(east_temp_lm)
```
```{r}
# Linear regression for chlorides in outlet
outlet_chlr_lm <- lm(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=outlet)
summary(outlet_chlr_lm)
```
```{r}
outlet_ph_lm <- lm(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=outlet)
summary(outlet_ph_lm)
```
```{r}
outlet_temp_lm <- lm(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=outlet)
summary(outlet_temp_lm)
```
```{r}
west_chlr_lm <- lm(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=west)
summary(west_chlr_lm)
```
```{r}
west_ph_lm <- lm(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=west)
summary(west_ph_lm)
```
```{r}
west_temp_lm <- lm(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip,data=west)
summary(west_temp_lm)
```

Looking at the same regression but with decision trees
```{r}
# Decision tree regression
library(rpart)
library(rpart.plot)
library(dplyr)
```

```{r}
# East analysis
east$id <- 1:nrow(east)
east_tree_train <- east %>% dplyr::sample_frac(0.7)
east_tree_test <- dplyr::anti_join(east,east_tree_train,by="id")

# Create decision tree model for chlorides for east sensor
east_chlr_tree <- rpart(Chloride_Water_EXO ~ se_precip + sw_precip + ne_precip + nw_precip, data=east_tree_train, method="anova")

# Create decision tree model for pH for east sensor
east_ph_tree <- rpart(pH_Water_EXO ~ se_precip + sw_precip + ne_precip + nw_precip,data=east_tree_train, method="anova")

# Create decision tree model for water temperature for east sensor
east_temp_tree <- rpart(Temperature_Water_EXO ~ se_precip + sw_precip + ne_precip + nw_precip,data=east_tree_train,method="anova")
```

Now need to create the same models for the Outlet and West sensors

```{r}
outlet$id <- 1:nrow(outlet)
outlet_tree_train <- outlet %>% dplyr::sample_frac(0.7)
outlet_tree_test <- dplyr::anti_join(outlet,outlet_tree_train, by="id")

# Create decision tree model for chlorides for outlet sensor
outlet_chlr_tree <- rpart(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=outlet_tree_train,method="anova")

# Create decision tree model for pH for outlet sensor
outlet_ph_tree <- rpart(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=outlet_tree_train,method="anova")

# Create decision tree model for water temperature for outlet sensor
outlet_temp_tree <- rpart(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=outlet_tree_train,method="anova")
```

Same code chunk for West sensor

```{r}
west$id <- 1:nrow(west)
west_tree_train <- west %>% dplyr::sample_frac(0.7)
west_tree_test <- dplyr::anti_join(west,west_tree_train,by="id")

# Create decision tree model for chlorides for west
west_chlr_tree <- rpart(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=west_tree_train,method="anova")

# Create decision tree model for pH for west
west_ph_tree <- rpart(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=west_tree_train, method="anova")

# Create decision tree model for water temperature for west
west_temp_tree <- rpart(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=west_tree_train, method="anova")
```

Now that I have all the decision tree models, I need to apply them to the test data

```{r}
# East

# East Chlorides
east_chlr_test <- predict(east_chlr_tree,east_tree_test,type="vector")
rpart.plot(east_chlr_tree, main="Decision Tree for East Chlorides")
head(east_chlr_test)

# East pH
east_ph_test <- predict(east_ph_tree, east_tree_test, type="vector")
rpart.plot(east_ph_tree, main="Decision Tree for East pH")
head(east_ph_test)

# East Temperature
east_temp_test <- predict(east_temp_tree, east_tree_test, type="vector")
rpart.plot(east_temp_tree, main="Decision Tree for East Temperature")
head(east_temp_test)
```

```{r}
# Outlet

# Outlet Chlorides
outlet_chlr_test <- predict(outlet_chlr_tree, outlet_tree_test, type="vector")
rpart.plot(outlet_chlr_tree,main="Decision Tree for Outlet Chlorides")
head(outlet_chlr_test)

# Outlet pH
outlet_ph_test <- predict(outlet_ph_tree,outlet_tree_test,type="vector")
rpart.plot(outlet_ph_tree,main="Decision Tree for Outlet pH")
head(outlet_ph_test)

# Outlet Temperature
outlet_temp_test <- predict(outlet_temp_tree,outlet_tree_test,type="vector")
rpart.plot(outlet_temp_tree,main="Decision Tree for Outlet Temperature")
head(outlet_temp_test)
```

```{r}
# West

# West Chlorides
west_chlr_test <- predict(west_chlr_tree,west_tree_test,type="vector")
rpart.plot(west_chlr_tree,main="Decision Tree for West Chlorides")
head(west_chlr_test)

# West pH
west_ph_test <- predict(west_ph_tree,west_tree_test,type="vector")
rpart.plot(west_ph_tree,main="Decision Tree for West pH")
head(west_ph_test)

# West Temperature
west_temp_test <- predict(west_temp_tree,west_tree_test,type="vector")
rpart.plot(west_temp_tree,main="Decision Tree for West Temperature")
head(west_temp_test)
```

The next model is a random forest model. We will look to compare the results from the
decision tree model to the random forest one. Luckily we should be able to use the same
data as the decision tree model.

```{r}
library(randomForest)

# Create random forest model for chlorides for east sensor
east_rf_chlr <- randomForest(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=east_tree_train,importance=TRUE,type="regression", na.action=na.omit)
# Plot one of the trees from the forest
getTree(east_rf_chlr,k=1,labelVar = TRUE)

# Create random forest model for pH for east sensor
east_rf_ph <- randomForest(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=east_tree_train,importance=TRUE,type="regression", na.action=na.omit)
# Plot one of the trees from the forest
getTree(east_rf_ph,k=1,labelVar=TRUE)

# Create random forest model for temperature for east sensor
east_rf_temp <- randomForest(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=east_tree_train, importance=TRUE,type="regression",na.action=na.omit)
# Plot one of the trees from the forest
getTree(east_rf_temp,k=1,labelVar=TRUE)

```

Create the models for the outlet sensor

```{r}
# Create random forest model for chlorides for outlet
outlet_rf_chlr <- randomForest(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=outlet_tree_train,importance=TRUE,type="regression",na.action=na.omit)
# Plot one of the trees
getTree(outlet_rf_chlr,k=1,labelVar=TRUE)

# Create random forest model for pH for outlet
outlet_rf_ph <- randomForest(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=outlet_tree_train,importance=TRUE,type="regression",na.action=na.omit)
# Plot one of the trees
getTree(outlet_rf_ph,k=1,labelVar=TRUE)

# Create random forest model for temperature for outlet
outlet_rf_temp <- randomForest(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=outlet_tree_train,importance=TRUE,type="regression",na.action=na.omit)
# Plot one of the trees
getTree(outlet_rf_temp, k=1, labelVar=TRUE)
```

Create the models for the east sensor

```{r}
# Create random forest model for chlorides for west
west_rf_chlr <- randomForest(Chloride_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=west_tree_train,importance=TRUE,type="regression",na.action=na.omit)
# Plot one tree from the forest
getTree(west_rf_chlr,k=1,labelVar=TRUE)

# Create random forest model for pH for west
west_rf_ph <- randomForest(pH_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=west_tree_train,importance=TRUE,type="regression",na.action=na.omit)
# Plot a tree from the forest
getTree(west_rf_ph, k=1,labelVar=TRUE)

# Create random forest model for temperature for west
west_rf_temp <- randomForest(Temperature_Water_EXO ~ sw_precip + se_precip + nw_precip + ne_precip, data=west_tree_train,importance=TRUE,type="regression",na.action=na.omit)
# Plot a tree
getTree(west_rf_temp,k=1,labelVar=TRUE)
```
For the last model, we will use KNN regression to predict the three variables

```{r}
# KNN for East
library(caTools)
library(class)
# Create new train and test sets due to knn function limitations
east_knn_train <- na.omit(east_tree_train)
east_knn_test <- na.omit(east_tree_test)

# Create predictors data frames so we only use the precipitation data
east_predictors_train <- data.frame(east_knn_train$nw_precip,east_knn_train$ne_precip,east_knn_train$sw_precip,east_knn_train$se_precip)
east_predictors_test <- data.frame(east_knn_test$nw_precip,east_knn_test$ne_precip,east_knn_test$sw_precip,east_knn_test$se_precip)

# East chlorides
east_knn_chlr <- knn(east_predictors_train,east_predictors_test,cl=east_knn_train$Chloride_Water_EXO,k=1)
# Create a data frame to house the results
east_knn_df <- data.frame(chlr_pred=east_knn_chlr,chlr_true=east_knn_test$Chloride_Water_EXO)

# East pH
east_knn_ph <- knn(east_predictors_train,east_predictors_test,cl=east_knn_train$pH_Water_EXO,k=1)
east_knn_df$ph_pred <- east_knn_ph
east_knn_df$ph_true <- east_knn_test$pH_Water_EXO

# East Temperature
east_knn_temp <- knn(east_predictors_train,east_predictors_test,cl=east_knn_train$Temperature_Water_EXO,k=1)
east_knn_df$temp_pred <- east_knn_temp
east_knn_df$temp_true <- east_knn_test$Temperature_Water_EXO
```

```{r}
# KNN for Outlet
library(caTools)
library(class)

# Create new train and test sets due to knn function limitation
outlet_knn_train <- na.omit(outlet_tree_train)
outlet_knn_test <- na.omit(outlet_tree_test)

# Create predictors data frames so we only use the precipitation data
outlet_predictors_train <- data.frame(outlet_knn_train$nw_precip,outlet_knn_train$ne_precip,outlet_knn_train$sw_precip,outlet_knn_train$se_precip)
outlet_predictors_test <- data.frame(outlet_knn_test$nw_precip,outlet_knn_test$ne_precip,outlet_knn_test$sw_precip,outlet_knn_test$se_precip)

# Outlet chlorides
outlet_knn_chlr <- knn(outlet_predictors_train,outlet_predictors_test,cl=outlet_knn_train$Chloride_Water_EXO,k=4)
# Create a data frame to house the results
outlet_knn_df <- data.frame(chlr_pred=outlet_knn_chlr,chlr_true=outlet_knn_test$Chloride_Water_EXO)

# Outlet pH
outlet_knn_ph <- knn(outlet_predictors_train,outlet_predictors_test,cl=outlet_knn_train$pH_Water_EXO,k=4)
outlet_knn_df$ph_pred <- outlet_knn_ph
outlet_knn_df$ph_true <- outlet_knn_test$pH_Water_EXO

# Outlet Temperature
outlet_knn_temp <- knn(outlet_predictors_train,outlet_predictors_test,cl=outlet_knn_train$Temperature_Water_EXO,k=4)
outlet_knn_df$temp_pred <- outlet_knn_temp
outlet_knn_df$temp_true <- outlet_knn_test$Temperature_Water_EXO
```

```{r}
# KNN for West
library(caTools)
library(class)

# Create new train and test sets due to knn function limitation
west_knn_train <- na.omit(west_tree_train)
west_knn_test <- na.omit(west_tree_test)

# Create predictors data frames so we only use the precipitation data
west_predictors_train <- data.frame(west_knn_train$nw_precip,west_knn_train$ne_precip,west_knn_train$sw_precip,west_knn_train$se_precip)
west_predictors_test <- data.frame(west_knn_test$nw_precip,west_knn_test$ne_precip,west_knn_test$sw_precip,west_knn_test$se_precip)

# West chlorides
west_knn_chlr <- knn(west_predictors_train,west_predictors_test,cl=west_knn_train$Chloride_Water_EXO,k=2)
# Create a data frame to house the results
west_knn_df <- data.frame(chlr_pred=west_knn_chlr,chlr_true=west_knn_test$Chloride_Water_EXO)

# West pH
west_knn_ph <- knn(west_predictors_train,west_predictors_test,cl=west_knn_train$pH_Water_EXO,k=2)
west_knn_df$ph_pred <- west_knn_ph
west_knn_df$ph_true <- west_knn_test$pH_Water_EXO

# West Temperature
west_knn_temp <- knn(west_predictors_train,west_predictors_test,cl=west_knn_train$Temperature_Water_EXO,k=2)
west_knn_df$temp_pred <- west_knn_temp
west_knn_df$temp_true <- west_knn_test$Temperature_Water_EXO
```


Need to predict the values in test
Plot residuals in predictions vs actual
  for i in 1:nrow(df):
    res = df_test[i] - prediction[i]
    res_list.append(res)
  plot(res)


Now to evaluate the models based on error between predicted and actual values

```{r}
# Predict using multiple regression for east

# East chlorides
east_pred_chlr_lm <- predict(east_chlr_lm,east_tree_test)
east_lm_df <- data.frame(east_pred_chlr_lm,east_tree_test$Chloride_Water_EXO)

# East pH
east_pred_ph_lm <- predict(east_ph_lm,east_tree_test)
east_lm_df$true_ph <- east_tree_test$pH_Water_EXO
east_lm_df$ph_pred <- east_pred_ph_lm

# East Temperature
east_pred_temp <- predict(east_temp_lm,east_tree_test)
east_lm_df$true_temp <- east_tree_test$Temperature_Water_EXO
east_lm_df$temp_pred <- east_pred_temp

# Get rid on any na values
east_lm_df <- na.omit(east_lm_df)

east_chlr_lm_mse <- mean((east_lm_df$east_pred_chlr_lm - east_lm_df$east_tree_test.Chloride_Water_EXO)^2)
east_ph_lm_mse <- mean((east_lm_df$ph_pred - east_lm_df$true_ph)^2)
east_temp_lm_mse <- mean((east_lm_df$temp_pred - east_lm_df$true_temp)^2)

# See the mse values
east_chlr_lm_mse
east_ph_lm_mse
east_temp_lm_mse
```

```{r}
# Predict using multiple regression for outlet

# Outlet chlorides
outlet_pred_chlr_lm <- predict(outlet_chlr_lm,outlet_tree_test)
outlet_lm_df <- data.frame(outlet_pred_chlr_lm,outlet_tree_test$Chloride_Water_EXO)

# East pH
outlet_pred_ph_lm <- predict(outlet_ph_lm,outlet_tree_test)
outlet_lm_df$true_ph <- outlet_tree_test$pH_Water_EXO
outlet_lm_df$ph_pred <- outlet_pred_ph_lm

# East Temperature
outlet_pred_temp <- predict(outlet_temp_lm,outlet_tree_test)
outlet_lm_df$true_temp <- outlet_tree_test$Temperature_Water_EXO
outlet_lm_df$temp_pred <- outlet_pred_temp

# Get rid on any na values
outlet_lm_df <- na.omit(outlet_lm_df)
# Calculate mse values
outlet_chlr_lm_mse <- mean((outlet_lm_df$outlet_pred_chlr_lm - outlet_lm_df$outlet_tree_test.Chloride_Water_EXO)^2)
outlet_ph_lm_mse <- mean((outlet_lm_df$ph_pred - outlet_lm_df$true_ph)^2)
outlet_temp_lm_mse <- mean((outlet_lm_df$temp_pred - outlet_lm_df$true_temp)^2)

# See the mse values
outlet_chlr_lm_mse
outlet_ph_lm_mse
outlet_temp_lm_mse
```

```{r}
# Predict using multiple regression for west

# Outlet chlorides
west_pred_chlr_lm <- predict(west_chlr_lm,west_tree_test)
west_lm_df <- data.frame(west_pred_chlr_lm,west_tree_test$Chloride_Water_EXO)

# East pH
west_pred_ph_lm <- predict(west_ph_lm,west_tree_test)
west_lm_df$true_ph <- west_tree_test$pH_Water_EXO
west_lm_df$ph_pred <- west_pred_ph_lm

# East Temperature
west_pred_temp <- predict(west_temp_lm,west_tree_test)
west_lm_df$true_temp <- west_tree_test$Temperature_Water_EXO
west_lm_df$temp_pred <- west_pred_temp

# Get rid on any na values
west_lm_df <- na.omit(west_lm_df)
# Calculate mse values
west_chlr_lm_mse <- mean((west_lm_df$west_pred_chlr_lm - west_lm_df$west_tree_test.Chloride_Water_EXO)^2)
west_ph_lm_mse <- mean((west_lm_df$ph_pred - west_lm_df$true_ph)^2)
west_temp_lm_mse <- mean((west_lm_df$temp_pred - west_lm_df$true_temp)^2)

# See the mse values
west_chlr_lm_mse
west_ph_lm_mse
west_temp_lm_mse
```

Now need to do the same process for decision tree models
```{r}
# Decision tree analysis for East
# East chlorides
east_pred_chlr_dt <- predict(east_chlr_tree,east_tree_test)
east_dt_df <- data.frame(pred_chlr=east_pred_chlr_dt, true_chlr=east_tree_test$Chloride_Water_EXO)

# East pH
east_pred_ph_dt <- predict(east_ph_tree,east_tree_test)
east_dt_df$pred_ph <- east_pred_ph_dt
east_dt_df$true_ph <- east_tree_test$pH_Water_EXO

# East temperature
east_pred_temp_dt <- predict(east_temp_tree,east_tree_test)
east_dt_df$pred_temp <- east_pred_temp_dt
east_dt_df$true_temp <- east_tree_test$Temperature_Water_EXO

east_dt_df <- na.omit(east_dt_df)

# Calculate the mse
east_chlr_dt_mse <- mean((east_dt_df$pred_chlr - east_dt_df$true_chlr)^2)
east_ph_dt_mse <- mean((east_dt_df$pred_ph - east_dt_df$true_ph)^2)
east_temp_dt_mse <- mean((east_dt_df$pred_temp - east_dt_df$true_temp)^2)

# See the mse values
east_chlr_dt_mse
east_ph_dt_mse
east_temp_dt_mse
```

```{r}
# Decision tree analysis for Outlet
# Outlet chlorides
outlet_pred_chlr_dt <- predict(outlet_chlr_tree,outlet_tree_test)
outlet_dt_df <- data.frame(pred_chlr=outlet_pred_chlr_dt, true_chlr=outlet_tree_test$Chloride_Water_EXO)

# Outlet pH
outlet_pred_ph_dt <- predict(outlet_ph_tree,outlet_tree_test)
outlet_dt_df$pred_ph <- outlet_pred_ph_dt
outlet_dt_df$true_ph <- outlet_tree_test$pH_Water_EXO

# Outlet temperature
outlet_pred_temp_dt <- predict(outlet_temp_tree,outlet_tree_test)
outlet_dt_df$pred_temp <- outlet_pred_temp_dt
outlet_dt_df$true_temp <- outlet_tree_test$Temperature_Water_EXO

outlet_dt_df <- na.omit(outlet_dt_df)

# Calculate the mse
outlet_chlr_dt_mse <- mean((outlet_dt_df$pred_chlr - outlet_dt_df$true_chlr)^2)
outlet_ph_dt_mse <- mean((outlet_dt_df$pred_ph - outlet_dt_df$true_ph)^2)
outlet_temp_dt_mse <- mean((outlet_dt_df$pred_temp - outlet_dt_df$true_temp)^2)

# See the mse values
outlet_chlr_dt_mse
outlet_ph_dt_mse
outlet_temp_dt_mse
```


```{r}
# Decision tree analysis for West
# West chlorides
west_pred_chlr_dt <- predict(west_chlr_tree,west_tree_test)
west_dt_df <- data.frame(pred_chlr=west_pred_chlr_dt, true_chlr=west_tree_test$Chloride_Water_EXO)

# West pH
west_pred_ph_dt <- predict(west_ph_tree,west_tree_test)
west_dt_df$pred_ph <- west_pred_ph_dt
west_dt_df$true_ph <- west_tree_test$pH_Water_EXO

# West temperature
west_pred_temp_dt <- predict(west_temp_tree,west_tree_test)
west_dt_df$pred_temp <- west_pred_temp_dt
west_dt_df$true_temp <- west_tree_test$Temperature_Water_EXO

west_dt_df <- na.omit(west_dt_df)

# Calculate the mse
west_chlr_dt_mse <- mean((west_dt_df$pred_chlr - west_dt_df$true_chlr)^2)
west_ph_dt_mse <- mean((west_dt_df$pred_ph - west_dt_df$true_ph)^2)
west_temp_dt_mse <- mean((west_dt_df$pred_temp - west_dt_df$true_temp)^2)

# See the mse values
west_chlr_dt_mse
west_ph_dt_mse
west_temp_dt_mse
```

Now to random forest analysis

```{r}
# Random forest analysis for East

# East Chlorides
east_pred_chlr_rf <- predict(east_rf_chlr,east_tree_test)
east_rf_df <- data.frame(pred_chlr=east_pred_chlr_rf,true_chlr = east_tree_test$Chloride_Water_EXO)

# East pH
east_pred_ph_rf <- predict(east_rf_ph,east_tree_test)
east_rf_df$pred_ph <- east_pred_ph_rf
east_rf_df$true_ph <- east_tree_test$pH_Water_EXO

# East Temperature
east_pred_temp_rf <- predict(east_rf_temp,east_tree_test)
east_rf_df$pred_temp <- east_pred_temp_rf
east_rf_df$true_temp <- east_tree_test$Temperature_Water_EXO

east_rf_df <- na.omit(east_rf_df)

# Calculate the mse
east_chlr_rf_mse <- mean((east_rf_df$pred_chlr - east_rf_df$true_chlr)^2)
east_ph_rf_mse <- mean((east_rf_df$pred_ph - east_rf_df$true_ph)^2)
east_temp_rf_mse <- mean((east_rf_df$pred_temp - east_rf_df$true_temp)^2)

# See the mse values
east_chlr_rf_mse
east_ph_rf_mse
east_temp_rf_mse
```
```{r}
# Random forest analysis for Outlet

# Outlet Chlorides
outlet_pred_chlr_rf <- predict(outlet_rf_chlr,outlet_tree_test)
outlet_rf_df <- data.frame(pred_chlr=outlet_pred_chlr_rf,true_chlr = outlet_tree_test$Chloride_Water_EXO)

# Outlet pH
outlet_pred_ph_rf <- predict(outlet_rf_ph,outlet_tree_test)
outlet_rf_df$pred_ph <- outlet_pred_ph_rf
outlet_rf_df$true_ph <- outlet_tree_test$pH_Water_EXO

# Outlet Temperature
outlet_pred_temp_rf <- predict(outlet_rf_temp,outlet_tree_test)
outlet_rf_df$pred_temp <- outlet_pred_temp_rf
outlet_rf_df$true_temp <- outlet_tree_test$Temperature_Water_EXO

outlet_rf_df <- na.omit(outlet_rf_df)

# Calculate the mse
outlet_chlr_rf_mse <- mean((outlet_rf_df$pred_chlr - outlet_rf_df$true_chlr)^2)
outlet_ph_rf_mse <- mean((outlet_rf_df$pred_ph - outlet_rf_df$true_ph)^2)
outlet_temp_rf_mse <- mean((outlet_rf_df$pred_temp - outlet_rf_df$true_temp)^2)

# See the mse values
outlet_chlr_rf_mse
outlet_ph_rf_mse
outlet_temp_rf_mse
```


```{r}
# Random forest analysis for West

# West Chlorides
west_pred_chlr_rf <- predict(west_rf_chlr,west_tree_test)
west_rf_df <- data.frame(pred_chlr=west_pred_chlr_rf,true_chlr = west_tree_test$Chloride_Water_EXO)

# West pH
west_pred_ph_rf <- predict(west_rf_ph,west_tree_test)
west_rf_df$pred_ph <- west_pred_ph_rf
west_rf_df$true_ph <- west_tree_test$pH_Water_EXO

# West Temperature
west_pred_temp_rf <- predict(west_rf_temp,west_tree_test)
west_rf_df$pred_temp <- west_pred_temp_rf
west_rf_df$true_temp <- west_tree_test$Temperature_Water_EXO

west_rf_df <- na.omit(west_rf_df)

# Calculate the mse
west_chlr_rf_mse <- mean((west_rf_df$pred_chlr - west_rf_df$true_chlr)^2)
west_ph_rf_mse <- mean((west_rf_df$pred_ph - west_rf_df$true_ph)^2)
west_temp_rf_mse <- mean((west_rf_df$pred_temp - west_rf_df$true_temp)^2)

# See the mse values
west_chlr_rf_mse
west_ph_rf_mse
west_temp_rf_mse
```
KNN Analysis

```{r}
# Knn analysis for East

# East chlorides
east_chlr_knn_mse <- mean((as.numeric(as.character(east_knn_df$chlr_pred)) - as.numeric(as.character(east_knn_df$chlr_true)))^2)
# East pH
east_ph_knn_mse <- mean((as.numeric(as.character(east_knn_df$ph_pred)) - as.numeric(as.character(east_knn_df$ph_true)))^2)
# East temperature
east_temp_knn_mse <- mean((as.numeric(as.character(east_knn_df$temp_pred)) - as.numeric(as.character(east_knn_df$temp_true)))^2)

# See the mse
east_chlr_knn_mse
east_ph_knn_mse
east_temp_knn_mse
```
```{r}
# Knn analysis for Outlet

# Outlet chlorides
outlet_chlr_knn_mse <- mean((as.numeric(as.character(outlet_knn_df$chlr_pred)) - as.numeric(as.character(outlet_knn_df$chlr_true)))^2)
# Outlet pH
outlet_ph_knn_mse <- mean((as.numeric(as.character(outlet_knn_df$ph_pred)) - as.numeric(as.character(outlet_knn_df$ph_true)))^2)
# East temperature
outlet_temp_knn_mse <- mean((as.numeric(as.character(outlet_knn_df$temp_pred)) - as.numeric(as.character(outlet_knn_df$temp_true)))^2)

# See the mse
outlet_chlr_knn_mse
outlet_ph_knn_mse
outlet_temp_knn_mse
```
```{r}
# Knn analysis for West

# West chlorides
west_chlr_knn_mse <- mean((as.numeric(as.character(west_knn_df$chlr_pred)) - as.numeric(as.character(west_knn_df$chlr_true)))^2)
# East pH
west_ph_knn_mse <- mean((as.numeric(as.character(west_knn_df$ph_pred)) - as.numeric(as.character(west_knn_df$ph_true)))^2)
# East temperature
west_temp_knn_mse <- mean((as.numeric(as.character(west_knn_df$temp_pred)) - as.numeric(as.character(west_knn_df$temp_true)))^2)

# See the mse
west_chlr_knn_mse
west_ph_knn_mse
west_temp_knn_mse
```

